Leadership across cultures and contexts: an exploratory analysis

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Editing for length

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::word_document2:
    number_sections: true
bibliography: Leader2.bib
csl: apa-6th-edition.csl
---

@@

<!-- ORIGINAL VERSION -->

@garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.

The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967]. The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions. The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers. See @garfield_evolutionary_2019 for additional details on the ethnographic sample.

ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*). See the *Coding example* section of the supplementary information (SI).

Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.

<!-- SHORTENDED VERSION -->

<!-- @garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. This database used the 60-culture Probability Sample Files (PSF) subsample [@naroll_proposed_1967] to capture a sample of the ethnographic record of leadership across diverse cultures and contexts. We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership. -->

<!-- We operationalized leader qualities and functions from the ethnographic text records in the database using an interative process based on multiple passes through many paragraphs, i.e., these operationalizations were not developed *a priori*. The final list of variables is given in Table S\@ref(tab:qualfuncvars) and KS and ZG coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions. -->

<!-- In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori*. For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)). Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny). The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs. have wives). -->

<!-- ZG and KS also coded all text records for group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. Finally, the sex of each leader was coded as male, female, both, or unknown. -->

<!-- ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*). See the *Coding example* section of the supplementary information (SI).  -->

<!-- <!-- ADD OTHER TEXT FROM ORIGINAL SECTION TO SI ?? --> -->

<!-- See @garfield_evolutionary_2019 and the SI for additional details on the ethnographic sample. -->

<!-- Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy. -->

@@

Our first goal was to assess if discovery of evidence for each of our leadership dimensions was biased by date of publication, total pages of ethnography for a culture, and sex of the ethnographer(s).

Our second goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions. We computed the proportion of text records supporting each dimension, and their 95% confidence intervals (CIs), using intercept-only logistic mixed effects models with random intercepts for document authors nested within cultures, fit with the lme4 package [@bates2015]. We then computed the proportion of cultures with at least one text record with evidence for each dimension, with 95% CIs computed using a cluster bootstrap.

Our third goal was to assess the extent to which each of the dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex. We again fit logistic mixed effects models using the lme4 package, with each leadership dimension as an outcome, all four factors entered as predictors, and author nested within culture as random intercepts. We also wanted to investigate which words were associated with particular leadership dimensions, or which dimensions predicted evidence for other dimensions. For these analyses, we often used elasticnet regression from the glmnet package [@friedman_regularization_2010]. Elasticnet models are penalized regression models that are useful when the number of predictors is large relative to the number of observations. We used the "lasso" penalty ($\alpha=1$), which will often set many coefficients to 0, thereby selecting the most important predictors.

@@

>>>>

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2:
    number_sections: true
bibliography: Leader2.bib
csl: apa-6th-edition.csl
---

@@

We use the database of `r nrow(all_data)` ethnographic paragraphs (termed text records) on leadership, described in @garfield_evolutionary_2019 (https://doi.org/10.5281/zenodo.2541999), to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership. The text records were extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) [@naroll_proposed_1967]. The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions.

The function and quality variables were not operationalized *a priori*. Instead, we used an iterative process whereby ZG and KS made multiple passes through the database, identifying leader functions and qualities (see Table S\@ref(tab:qualfuncvars)). In contrast, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature [@garfield_review_2019] (see Table S\@ref(tab:costbenefitvars)). ZG and KS then coded all text records for the presence (1) or absence (0) of evidence for each variable, which we henceforth refer to as leadership dimensions. Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny). The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs. have wives). ZG then created a measure of group context (see Table \@ref(tab:groupcontext)). ZG and KS coded all text records for group context. Finally, the sex of each leader was coded as male, female, both, or unknown.

```{r groupcontext}
gc <- read_tsv('group context table.tsv')
kable(gc, caption="Operationalization of the group context variable.")
```

ZG and KS discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus database. See the *Coding example* section of the supplementary information (SI). Finally, each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy.

@@

Our first goal was to assess if discovery of evidence for each of our leadership dimensions was biased by date of publication, total pages of ethnography for a culture, and sex of the ethnographer(s). Our second goal was to assess the proportion of text records and cultures that provided evidence for each of our coded dimensions. We computed the proportion of text records supporting each dimension, and their 95% confidence intervals (CIs), using intercept-only logistic mixed effects models with random intercepts for document authors nested within cultures, fit with the lme4 package [@bates2015]. We then computed the proportion of cultures with at least one text record with evidence for each dimension, with 95% CIs computed using a cluster bootstrap.

Our third goal was to assess the extent to which each of the dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex. We again fit logistic mixed effects models using the lme4 package, with each leadership dimension as an outcome, all four factors entered as predictors, and author nested within culture as random intercepts. To investigate which words were associated with particular leadership dimensions, or which dimensions predicted evidence for other dimensions, we used elasticnet regression from the glmnet package [@friedman_regularization_2010]. Elasticnet models are penalized regression models that are useful when the number of predictors is large relative to the number of observations. We used the "lasso" penalty ($\alpha=1$), which will often set many coefficients to 0, thereby selecting the most important predictors.

@@

## Variable coding details
